---
title: "Final Project"
author: "Akash Agrawal Bejarano"
date: "1/17/2023"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 1 

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
dataset <- read.csv("expenses.csv")

```


The dataset contains information regarding healthcare expenses with a total of 1338 observations. It has a total of 7 variables, 4 quantitative and 3 being categorical. The quantitative variables are Age, BMI (body mass index), children (how many children do you have), charges (cost). The categorical variables are Sex, Smoker, and Region. 

## 2

```{r}

```


## 3 

```{r}
fullmodel <- lm(charges ~ ., data = dataset)

```

## 4

## 5
```{r}
summary(fullmodel)
```

Adjusted R^2 value = 0.7494 

From the output summary above we can see that Age, BMI, children, smokeryes, region southeast, region southwest are all significant. while sexmale, region northwest are insignificant at the 0.05 significance level. 

## 7

Since sex and region both contained an insignificant effect on charges, we choose to remove these from the reduced model so that we only use significant predictors. 

```{r}
reducedmodel <- lm(charges ~ age + bmi + children + smoker, data = dataset)
summary(reducedmodel)
summary(fullmodel)
```























